Generating exact lattices in a WFST framework

### 2011-09-27 misc/papers/icassp12_lat/paper.tex
Improvements to ICASSP paper.. nearly finished.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@529 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

\name{ \em Daniel Povey$^1$, Mirko Hannemann$^{1,2}$,  \\
  \em {Gilles Boulianne}$^3$, {Luk\'{a}\v{s} Burget}$^4$, {Arnab Ghoshal}$^5$, {Milos Janda}$^2$, {Stefan Kombrink}$^2$, \\
  \em {Petr Motl\'{i}\v{c}ek}$^6$, {Yanmin Qian}$^7$, {Ngoc Thang Vu}$^8$, {Korbinian Riedhammer}$^9$, {Karel Vesel\'{y}}$^2$
    \thanks{Thanks here.. remember Sanjeev.} }

@@

\begin{figure}
\begin{center}
\includegraphics[height=2.1in,angle=270]{figures/acceptor_utterance.eps}
  \caption{Acceptor $U$ describing the acoustic scores of an utterance}
\label{fig:acceptor}
\end{center}
\end{figure}

@@

We will briefly describe another unique aspect of our
algorithm.  Determinization algorithms involve weighted subsets of states,
of the form
\begin{equation}
   S = \{ (s_1, w_1), (s_2, w_2), \ldots \} .
\end{equation}
Let this weighted subset, as it appears in a conventional determinization
algorithm with epsilon removal, be the {\em canonical representation} of a state.
A typical determinization algorithm would maintain a map from this representation
to a state index.  We define a {\em minimial representation} of a state
to be like the canonical representation, but only keeping states that
are either final, or have non-$\epsilon$ arcs out of them.  We maintain
a map from the minimal representation to the state index.  We can
show that this algorithm is still correct (it will tend to give output
with fewer states, i.e. more minimal).  As an optimization for speed, we also define
the {\em initial representation} to be the same type of subset, but prior
to following through the $\epsilon$ arcs, i.e. it only contains the states
that we reached by following non-$\epsilon$ arcs from a previous determinized
state.  We maintain a separate map from the initial representation to
the output state index; think of this as a ``lookaside buffer'' that helps us
avoid the expense of following $\epsilon$ arcs.

@@

We report experimental results on the Wall Street Journal database of
read speech.  
Our system is a standard mixture-of-Gaussians system trained on the SI-284
training data; we test on the November 1992 evaluation data.  
For these experiments we generate lattices with the bigram language model
supplied with the WSJ database, and rescore them with the trigram language model.

@@

Conclusions here [mention does not rely on word-pair assumption]

>>>>

\name{ \em Daniel Povey$^1$, Mirko Hannemann$^{1,2}$,  \\
  \em {Gilles Boulianne}$^3$, {Luk\'{a}\v{s} Burget}$^4$, {Arnab Ghoshal}$^5$, {Milos Janda}$^2$, {Stefan Kombrink}$^2$, \\
  \em {Petr Motl\'{i}\v{c}ek}$^6$, {Yanmin Qian}$^7$, {Ngoc Thang Vu}$^8$, {Korbinian Riedhammer}$^9$, {Karel Vesel\'{y}}$^2$
    \thanks{Thanks to Sanjeev Khudanpur for his help in preparing the paper}}

@@

\begin{figure}
\begin{center}
\includegraphics[height=2.1in,angle=270]{figures/acceptor_utterance.eps}
  \caption{Acceptor $U$ describing the acoustic scores of an utterance}
\vspace*{-0.2in}
\label{fig:acceptor}
\end{center}

\end{figure}

@@

We will briefly describe another unique aspect of our
algorithm.  Determinization algorithms involve weighted subsets of states,
of the form
\begin{equation}
   S = \{ (s_1, w_1), (s_2, w_2), \ldots \} .
\end{equation}
Let this weighted subset, as it appears in a conventional determinization
algorithm with epsilon removal, be the {\em canonical representation} of a state.
A typical determinization algorithm would maintain a map from this representation
to a state index.  We define a {\em minimial representation} of a state
to be like the canonical representation, but only keeping states that
are either final, or have non-$\epsilon$ arcs out of them.  We maintain
a map from the minimal representation to the state index.  We can
show that this algorithm is still correct (it will tend to give more
minimal output).   As an optimization for speed, we also define
the {\em initial representation} to be the same type of subset, but prior
to following through the $\epsilon$ arcs, i.e. it only contains the states
that we reached by following non-$\epsilon$ arcs from a previous determinized
state.  We maintain a separate map from the initial representation to
the output state index; think of this as a ``lookaside buffer'' that helps us
avoid the expense of following $\epsilon$ arcs.

@@

We report experimental results on the Wall Street Journal database of
read speech.  
Our system is a standard mixture-of-Gaussians system trained on the SI-284
training data; we test on the November 1992 evaluation data.  
For these experiments we generate lattices with the bigram language model
supplied with the WSJ database, and for rescoring experiments we
use the trigram language model.  The acoustic scale was $1/16$ for first-pass
decoding and $1/15$ for LM rescoring.

For simplicity, our results were all obtained using a decoder that
does not support a ``maximum active states'' option, so the only variables
to consider are the beam used in the Viterbi beam search, and the separate
beam $\alpha$ used for lattice generation.

\begin{figure}
\centering
   \includegraphics[width=0.9\columnwidth]{figures/latbeam.eps} 
 \vspace*{-0.1in}

   \caption{\vspace*{-0.2in} Varying lattice beam $\alpha$ (Viterbi pruning beam fixed at 15) }
  \label{fig:latbeam}
\end{figure}

Figure~\ref{fig:latbeam} shows how various quantities change as we vary
$\alpha$, with the Viterbi beam fixed at 15.  Note that we get all the improvement from
LM rescoring by increasing $\alpha$ to 4.  The time taken by our algorithm started to increase rapidly after about
$\alpha=8$, so a value of $\alpha$ anywhere between about 4 and 8 is sufficient for LM rescoring 
and still does not slow down decoding too much.
Note that out of vocabularly words (OOVs) provide a floor on the lattice oracle error rate.
Of the 333 test utterances, 87 contained an OOV word, yet only 93 sentences had oracle errors
with $\alpha=10$.  Figure~\ref{fig:viterbibeam} shows the effect of varying the
Viterbi decoding beam, while leaving $\alpha$ fixed at 7.

\begin{figure}
\centering
   \includegraphics[width=0.9\columnwidth]{figures/decodebeam.eps} 
 \vspace*{-0.1in}
    \caption{\vspace*{-0.2in} Varying Viterbi pruning beam (lattice beam $\alpha$ fixed at 7) }
  \label{fig:viterbibeam}
\end{figure}

@@

We have described a lattice generation method that is to our knowledge the first
efficient method not to rely on the word-pair assumption of~\cite{ney_word_graph}.
It includes an ingenious way of obtaining state-level alignment information
via determinization in a specially designed semiring.
